---
title: "Auto Mobile price Analysis Project - EDA In Depth"
author: 'author : Omar Soub'
date: "date : 23-05/2024"
output:
  html_document:
    toc: yes
    toc_float:
      collapsed: true
    
    
---



# *Introduction* 

* *On this chapter ,we will use R-Programming Language in order to conduct inferential analysis to analyse the data as much as we can to help decision making*

----------------------------------------------

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(GGally)
library(lubridate)
library(skimr)
library(readxl)
library(janitor)
library(gridExtra)
library(flextable)
library(knitr)
library(gridExtra)
library(patchwork)
library(dlookr)
library(rstatix)
library(ggstatsplot)
library(effectsize)
library(FSA)
library(ISLR)
library(repr)
library(gtsummary)
library(corrplot)
```


```{r message=FALSE, warning=FALSE, include=FALSE}
df<-read.csv("C:/Users/Omar/Desktop/Omar_Files/Python_Analysis/AutoMobile_Data_Analysis_2/new_cleaned_df.csv")
```

# *Data Set and Data types view*

----------------------------------------------

## *Data Set (1st five rows) :*

```{r echo=FALSE, message=FALSE, warning=FALSE}
df %>% head() %>% kable()
```

----------------------------------------------

## *Data types and some details using glimps function*

```{r echo=FALSE, message=FALSE, warning=FALSE}
df %>% glimpse()
```

----------------------------------------------

* *convert variables from character data type into factor*

```{r message=FALSE, warning=FALSE, include=FALSE}
df[sapply(df, is.character)] <- lapply(df[sapply(df, is.character)], 
                                                           as.factor)
```


```{r message=FALSE, warning=FALSE}

# df[sapply(df, is.character)] <- lapply(df[sapply(df, is.character)],as.factor)                                
```

----------------------------------------------

## *Data types and some details using skim function*

```{r echo=FALSE, message=FALSE, warning=FALSE}
df %>% skim()
```

-----------------------------------------------

###  *Define categorical variables*  

```{r message=FALSE, warning=FALSE, include=FALSE}
categorical_data<-df %>% select_if(is.factor)
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
categorical_data %>% colnames()

```

-----------------------------------------------

###  *Define numerical variables*  

```{r message=FALSE, warning=FALSE, include=FALSE}

numerical_data<-df %>% select_if(is.numeric) 
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
numerical_data %>% colnames()
```


-----------------------------------------------

# *Categorical variables visualization*


```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
( 
  ggplot(df, aes(y=body_style, fill=body_style))
    + geom_bar()
    + geom_text(
        aes(label=after_stat(count)), stat="count", nudge_y=0.125, va="bottom")+theme(legend.position = "none")
)+labs(x="",y="",title="Count of body_style variable")+
  
( 
  ggplot(df, aes(y=fuel_system, fill=fuel_system))
    + geom_bar()
    + geom_text(
        aes(label=after_stat(count)), stat="count", nudge_y=0.125, va="bottom")+theme(legend.position = "none")
)+labs(x="",y="",title="Count of fuel_system variable")+
( 
  ggplot(df, aes(y=num_of_cylinders, fill=num_of_cylinders))
    + geom_bar()
    + geom_text(
        aes(label=after_stat(count)), stat="count", nudge_y=0.125, va="bottom")+theme(legend.position = "none")
)+labs(x="",y="",title="Count of num_of_cylinders variable")
```
  


  

```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
( 
  ggplot(df, aes(y=num_of_doors, fill=num_of_doors))
    + geom_bar()
    + geom_text(
        aes(label=after_stat(count)), stat="count", nudge_y=0.125, va="bottom")+theme(legend.position = "none")
)+labs(x="",y="",title="Count of num_of_doors variable")+
  
( 
  ggplot(df, aes(y=drive_wheels, fill=drive_wheels))
    + geom_bar()
    + geom_text(
        aes(label=after_stat(count)), stat="count", nudge_y=0.125, va="bottom")+theme(legend.position = "none")
)+labs(x="",y="",title="Count of drive_wheels variable")+
  
( 
  ggplot(df, aes(y=engine_location, fill=engine_location))
    + geom_bar()
    + geom_text(
        aes(label=after_stat(count)), stat="count", nudge_y=0.125, va="bottom")+theme(legend.position = "none")
)+labs(x="",y="",title="Count of engine_location variable")  
```

-----------------------------------------------

# *Numerical variables visualization*


```{r echo=FALSE, fig.height=3, fig.width=10, message=FALSE, warning=FALSE}

ggplot(numerical_data, aes(x = price,y=after_stat(density))) + geom_histogram(fill = 'blue', color = 'black',alpha=.5) + labs(x="",y="",title="Distribution of price variable")+geom_density(color = "green", linewidth = 1)+
  
ggplot(numerical_data, aes(x = width,y=after_stat(density)))+geom_histogram(fill = 'blue', color = 'black',alpha=.5) + labs(x="",y="",title="Distribution of width variable")+geom_density(color = "green", linewidth = 1)+

ggplot(numerical_data, aes(x = height,y=after_stat(density)))+geom_histogram(fill = 'blue', color = 'black',alpha=.5) + labs(x="",y="",title="Distribution of height variable")+geom_density(color = "green", linewidth = 1)

  
```


```{r echo=FALSE, fig.height=3, fig.width=10, message=FALSE, warning=FALSE}

ggplot(numerical_data, aes(x = curb_weight,y=after_stat(density)))+geom_histogram(fill = 'blue', color = 'black',alpha=.5) + labs(x="",y="",title="Distribution of curb_weight variable")+geom_density(color = "green", linewidth = 1)+
  
ggplot(numerical_data, aes(x = engine_size,y=after_stat(density)))+geom_histogram(fill = 'blue', color = 'black',alpha=.5) + labs(x="",y="",title="Distribution of engine_size variable")+geom_density(color = "green", linewidth = 1)+
  
ggplot(numerical_data, aes(x = peak_rpm,y=after_stat(density)))+geom_histogram(fill = 'blue', color = 'black',alpha=.5) + labs(x="",y="",title="Distribution of peak_rpm variable")+geom_density(color = "green", linewidth = 1)
    
```




```{r echo=FALSE, fig.height=3, fig.width=10, message=FALSE, warning=FALSE}

ggplot(numerical_data, aes(x = length,y=after_stat(density)))+geom_histogram(fill = 'blue', color = 'black',alpha=.5) + labs(x="",y="",title="Distribution of length variable")+geom_density(color = "green", linewidth = 1)+
  
ggplot(numerical_data, aes(x = city_mpg,y=after_stat(density)))+geom_histogram(fill = 'blue', color = 'black',alpha=.5) + labs(x="",y="",title="Distribution of city_mpg variable")+geom_density(color = "green", linewidth = 1)+


ggplot(numerical_data, aes(x = highway_mpg,y=after_stat(density)))+geom_histogram(fill = 'blue', color = 'black',alpha=.5) + labs(x="",y="",title="Distribution of highway_mpg variable")+geom_density(color = "green", linewidth = 1) 
```

-----------------------------------------------

# *Visualizing The Relation between numerical variables :*


```{r echo=FALSE, fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
ggpairs(numerical_data)
```

-----------------------------------------------

# *Visualizing The Relation between The Price variable and other numerical variables :*

```{r fig.height=5, fig.width=10, message=FALSE, warning=FALSE, include=FALSE}
plot1 <- ggplot(df, aes(x = length, y = price)) + geom_point() + ggtitle("length vs price")+geom_smooth()
plot2 <- ggplot(df, aes(x = width, y = price)) + geom_point() + ggtitle("width vs price")+geom_smooth()
plot3 <- ggplot(df, aes(x = height, y = price)) + geom_point() + ggtitle("height vs price")+geom_smooth()
plot4 <- ggplot(df, aes(x = curb_weight, y = price)) + geom_point() + ggtitle("curb_weight vs price")+geom_smooth()
plot5 <- ggplot(df, aes(x = engine_size, y = price)) + geom_point() + ggtitle("engine_size vs price")+geom_smooth()
plot6 <- ggplot(df, aes(x = peak_rpm, y = price)) + geom_point() + ggtitle("peak_rpm vs price")+geom_smooth()
plot7 <- ggplot(df, aes(x = city_mpg, y = price)) + geom_point() + ggtitle("city_mpg vs price")+geom_smooth()
plot8 <- ggplot(df, aes(x = highway_mpg, y = price)) + geom_point() + ggtitle("highway_mpg vs price")+geom_smooth()
```


```{r echo=FALSE, fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
grid.arrange(plot1, plot2, plot3,plot4,plot5,plot6,plot7,plot8, ncol = 3)
```

> *We can see from the above charts that the relationship between the variables is monotonic (consistently increasing or decreasing but not necessarily linear)*

-----------------------------------------------

# *Visualizing The correlation between The numerical variables :*

* *as the data is not normally distributed we will apply spearman methods*

```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}

corrplot(cor(numerical_data,method = "spearman"),method="ellipse",order = "AOE",addCoef.col="black",
         cex.lab=.5, cex.axis=.5)
```


-----------------------------------------------

# *E. inferential analysis*


## *E.1: Price by number of doors*

* *E.1.1 : Price by number of doors-Basic description*


```{r echo=FALSE, message=FALSE, warning=FALSE}
df %>% group_by(num_of_doors) %>% summarise(
  count = n(),
  mean = mean(price),
  sd = sd(price),
  min = min(price),
  Q1 = quantile(price, 0.25),
  median = median(price),
  Q3 = quantile(price, 0.75),
  max = max(price)
) %>% kable()

```


```{r echo=FALSE, fig.height=3, fig.width=10, message=FALSE, warning=FALSE}
ggplot(df, aes(x = price,y=after_stat(density),fill=num_of_doors)) + geom_histogram( color = 'black',alpha=.3) + labs(x="",y="",title="Comparison of price variable Distribution bwteen num_of_doors ")+geom_density(alpha=.3)+facet_wrap(~num_of_doors)+theme(legend.position = "none")
```



* *E.1.2 : Applying wilcox.test (for variables with two groups) to find out if there are significant differences in Price Distribution among the Auto Mobiles based on number of doors groups :*
```{r echo=FALSE, message=FALSE, warning=FALSE}
wilcox.test(price ~ num_of_doors, data = df)
```
> *As the The p-value is more than 0.05 , we will accept the null hypothesis,indicating that there are no significant differences in Price Distribution among the Auto Mobiles based on number of doors groups.*




* *E.1.3 :Visualizing The Statistical test*

* *we will use Mann-Whitney U Test -for features that have 2 categories : equivelent to wilcox.test() wilcox rank sum test with continuity correction,better than wilcox sighned rank exct test(that used when variables are paired)*

```{r message=FALSE, warning=FALSE, include=FALSE}
ggstatsplot_customized_theme<-theme(text = element_text(size = 20),
                               plot.subtitle = element_text(size = 20),
                               legend.title = element_text(size = 30),
                               legend.text = element_text(size = 20))
```



```{r echo=FALSE, fig.height=3, fig.width=10, message=FALSE, warning=FALSE}
df %>% ggbetweenstats(x=num_of_doors,y=price,type = "nonparamertic",conf.level = 0.95,
                      centrality.label.args = list(size = 8, nudge_x = 0.6,
                                                   segment.linetype = 4,min.segment.length = 0))+labs(x="",y="Price")+
  ggstatsplot_customized_theme 
  
```



## *E.2: Price by drive_wheels*

* *E.2.1 : Price by drive_wheels-Basic description*

```{r echo=FALSE, message=FALSE, warning=FALSE}
df %>% group_by(drive_wheels ) %>% summarise(
  count = n(),
  mean = mean(price),
  sd = sd(price),
  min = min(price),
  Q1 = quantile(price, 0.25),
  median = median(price),
  Q3 = quantile(price, 0.75),
  max = max(price)
) %>% kable()

```


```{r echo=FALSE, fig.height=3, fig.width=10, message=FALSE, warning=FALSE}
ggplot(df, aes(x = price,y=after_stat(density),fill=drive_wheels)) + geom_histogram( color = 'black',alpha=.3) + labs(x="",y="",title="Comparison of price variable Distribution bwteen drive_wheels ")+geom_density(alpha=.3)+facet_wrap(~drive_wheels)+theme(legend.position = "none")
```


* *E.2.2 :Applying kruskal.test (for variables with more than two groups) to find out if there are significant differences in Price Distribution among the Auto Mobiles based on drive wheels groups :*
```{r echo=FALSE, message=FALSE, warning=FALSE}
kruskal.test(price ~ drive_wheels , data = df)
```

> *As the The p-value is less than 0.05 , we will reject the null hypothesis,indicating that there are significant differences in Price Distribution among the Auto Mobiles based on drive wheels groups.*


* *E.2.3 :Visualizing The Statistical test*


```{r echo=FALSE, fig.height=3, fig.width=10, message=FALSE, warning=FALSE}
df %>% ggbetweenstats(x=drive_wheels,y=price,type = "nonparamertic",conf.level = 0.95,pairwise.display = "none",
                      centrality.label.args = list(size = 8, nudge_x = 0.6,segment.linetype = 4,min.segment.length = 0),
                      ggsignif.args = list(textsize = 5, tip_length = 0.01, na.rm = TRUE),
                      p.adjust.method = "bonferroni")+
  labs(x="",y="Price")+
  ggstatsplot_customized_theme 
```




* *E.2.4: To find out this difference in details: we will apply pairwise.wilcox.test*

```{r echo=FALSE, message=FALSE, warning=FALSE}
pairwise.wilcox.test(df$price, df$drive_wheels, p.adjust.method = "bonferroni") 
```
> *As we can see there is significant difference in Price Distributions when we compare rwd group with the other two groups (4wd,fwd) ,as The p-value adjusted is less than 0.05, but the Price Distributions has no significant difference between groups (4wd,fwd) as The p-value adjusted is more than 0.05 *



## *E.3: Price by body_style*

* *E.3.1 : Price by body_style-Basic description*

```{r echo=FALSE, message=FALSE, warning=FALSE}
df %>% group_by(body_style ) %>% summarise(
  count = n(),
  mean = mean(price),
  sd = sd(price),
  min = min(price),
  Q1 = quantile(price, 0.25),
  median = median(price),
  Q3 = quantile(price, 0.75),
  max = max(price)
) %>% kable()

```


```{r echo=FALSE, fig.height=3, fig.width=10, message=FALSE, warning=FALSE}
ggplot(df, aes(x = price,y=after_stat(density),fill=body_style)) + geom_histogram( color = 'black',alpha=.3) + labs(x="",y="",title="Comparison of price variable Distribution bwteen body_style ")+geom_density(alpha=.3)+facet_wrap(~body_style)+theme(legend.position = "none")
```


* *E.3.2 :Applying kruskal.test (for variables with more than two groups) to find out if there are significant differences in Price Distribution among the Auto Mobiles based on body_style groups :*

```{r echo=FALSE, message=FALSE, warning=FALSE}
kruskal.test(price ~ body_style , data = df)
```


> *As the The p-value is less than 0.05 , we will reject the null hypothesis,indicating that there are significant differences in Price among the Auto Mobiles based on body style groups.*


* *E.3.3 :Visualizing The Statistical test*


```{r echo=FALSE, fig.height=3, fig.width=10, message=FALSE, warning=FALSE}
df %>% ggbetweenstats(x=body_style,y=price,type = "nonparamertic",conf.level = 0.95,pairwise.display = "none",
                      centrality.label.args = list(size = 8, nudge_x = 0.6,segment.linetype = 4,min.segment.length = 0),
                      ggsignif.args = list(textsize = 5, tip_length = 0.01, na.rm = TRUE),
                      p.adjust.method = "bonferroni")+
  labs(x="",y="Price")+
  ggstatsplot_customized_theme 
```




* *E.3.4: To find out this difference in details: we will apply pairwise.wilcox.test*


```{r echo=FALSE, message=FALSE, warning=FALSE}
pairwise.wilcox.test(df$price, df$body_style, p.adjust.method = "bonferroni")
```

> *As we can see there is significant difference in Price Distributions where the p-value adjusted is less than 0.05 when we compare :<br> 1.hatchback group with convertible group.<br> 2.sedan group with hatchback group.*




## *E.4: Price by body_style*

* *E.4.1 : Price by fuel_system-Basic description*

```{r echo=FALSE, message=FALSE, warning=FALSE}
df %>% group_by(fuel_system ) %>% summarise(
  count = n(),
  mean = mean(price),
  sd = sd(price),
  min = min(price),
  Q1 = quantile(price, 0.25),
  median = median(price),
  Q3 = quantile(price, 0.75),
  max = max(price)
) %>% kable()

```




```{r echo=FALSE, fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
ggplot(df, aes(x = price,y=after_stat(density),fill=fuel_system)) + geom_histogram( color = 'black',alpha=.3) + labs(x="",y="",title="Comparison of price variable Distribution bwteen fuel_system ")+geom_density(alpha=.3)+facet_wrap(~fuel_system)+theme(legend.position = "none")
```


* *E.4.2 :Applying kruskal.test (for variables with more than two groups) to find out if there are significant differences in Price Distribution among the Auto Mobiles based on fuel_system groups :*

```{r echo=FALSE, message=FALSE, warning=FALSE}
kruskal.test(price ~ fuel_system , data = df)
```
> *As the The p-value is less than 0.05 , we will reject the null hypothesis,indicating that there are significant differences in Price Distribution among the Auto Mobiles based on fuel_system groups.*


* *E.4.3 :Visualizing The Statistical test*


```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
df %>% ggbetweenstats(x=fuel_system,y=price,type = "nonparamertic",conf.level = 0.95,pairwise.display = "none",
                      centrality.label.args = list(size = 8, nudge_x = 0.6,segment.linetype = 4,min.segment.length = 0),
                      ggsignif.args = list(textsize = 5, tip_length = 0.01, na.rm = TRUE),
                      p.adjust.method = "bonferroni")+
  labs(x="",y="Price")+
  ggstatsplot_customized_theme 
```




* *E.4.4: To find out this difference in details: we will apply pairwise.wilcox.test*

```{r echo=FALSE, message=FALSE, warning=FALSE}
pairwise.wilcox.test(df$price, df$fuel_system, p.adjust.method = "bonferroni") 
```

> *As we can see there is significant difference in Price Distributions where the p-value adjusted is less than 0.05 when we compare :<br> 1. idi group with 1bbl and 2bbl groups.<br> 2. mpfi group with 1bbl and 2bbl groups.<br> 3. spdi group with 2bbl group.*




## *E.5: Price by num_of_cylinders*

* *E.5.1 : Price by num_of_cylinders-Basic description*

```{r echo=FALSE, message=FALSE, warning=FALSE}
df %>% group_by(num_of_cylinders ) %>% summarise(
  count = n(),
  mean = mean(price),
  sd = sd(price),
  min = min(price),
  Q1 = quantile(price, 0.25),
  median = median(price),
  Q3 = quantile(price, 0.75),
  max = max(price)
) %>% kable()

```




```{r echo=FALSE, fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
ggplot(df, aes(x = price,y=after_stat(density),fill=num_of_cylinders)) + geom_histogram( color = 'black',alpha=.3) + labs(x="",y="",title="Comparison of price variable Distribution bwteen num_of_cylinders ")+geom_density(alpha=.3)+facet_wrap(~num_of_cylinders)+theme(legend.position = "none")
```


* *E.5.2 :Applying kruskal.test (for variables with more than two groups) to find out if there are significant differences in Price Distribution among the Auto Mobiles based on fuel_system groups :*

```{r echo=FALSE, message=FALSE, warning=FALSE}
kruskal.test(price ~ num_of_cylinders , data = df)
```

> *As the The p-value is less than 0.05 , we will reject the null hypothesis,indicating that there are significant differences in Price Distribution among the Auto Mobiles based on num_of_cylinders groups.*


* *E.5.3 :Visualizing The Statistical test*

```{r echo=FALSE, fig.height=3, fig.width=10, message=FALSE, warning=FALSE}
df %>% ggbetweenstats(x=num_of_cylinders,y=price,type = "nonparamertic",conf.level = 0.95,pairwise.display = "none",
                      centrality.label.args = list(size = 8, nudge_x = 0.6,segment.linetype = 4,min.segment.length = 0),
                      ggsignif.args = list(textsize = 5, tip_length = 0.01, na.rm = TRUE),
                      p.adjust.method = "bonferroni")+
  labs(x="",y="Price")+
  ggstatsplot_customized_theme 
```



* *E.5.4: To find out this difference in details: we will apply pairwise.wilcox.test*

```{r echo=FALSE, message=FALSE, warning=FALSE}
pairwise.wilcox.test(df$price, df$num_of_cylinders, p.adjust.method = "bonferroni") 
```

> *As we can see there is significant difference in Price Distributions where the p-value adjusted is less than 0.05 when we compare :<br> 1. four group with five group.<br> 2. six group with four group.*

> *We will exclude cylinders Three and Twelve as there are no enough number of samples*

-----------------------------------------------

# *Final Conclusions:*

* *1. The number of doors (two doors vs. four doors) does not significantly affect the price distribution of the automobiles in the data set.*

* *2. The type of drive wheels (e.g., front-wheel drive, rear-wheel drive, or four-wheel drive) significantly affects the price distribution of the automobiles in the data set especially between groups (4wd,fwd) .*

* *3. The body style of automobiles (e.g., sedan, hatchback, SUV) significantly affects the price distribution of the automobiles in the data set especially between :<br> A. hatchback group with convertible group.<br>B. sedan group with hatchback group.*

* *4. The type of fuel system in automobiles (e.g., fuel injection, carburetor) significantly affects the price distribution of the automobiles in the data set especially between :<br> A. idi group with 1bbl and 2bbl groups.<br> B. mpfi group with 1bbl and 2bbl groups.<br> C. spdi group with 2bbl group.*

* *5. The number of cylinders in an automobile's engine significantly affects the price distribution of the automobiles in the data set especially between :<br> A. four group with five group.<br> B. six group with four group.*
